---
title: "Final Project: NCAA Men's March Madness"
format: html
editor: visual
---

## Loading Data

The data set used in this project is from the Tidy Tuesday GitHub, specifically from the week of 3-26-2024. The data set is titled "NCAA Men's March Madness", and it comes from Nishaan Amin's Kaggle data set and analysis "Bracketology: predicting March Madness". The data is from 2008-2024 for the men's teams. The year 2020 is not included because the tournament was cancelled due to Covid.

```{r}
#| message: false
load(here::here("data", "team_results.Rdata"))

library(tidyverse)
library(gtsummary)
library(ggplot2)
```

## Descriptive Statistics

@tbl-one shows the descriptive statistics of selected characteristics by number of championships won from 2008-2024 (excluding 2020).

```{r}
#| label: tbl-one
#| tbl-cap: "Men's March Madness Team Characteristics Stratified by Number of Championships Won from 2008-2024 (Excluding 2020)"

table1 <- tbl_summary(
	team_results,
	by = CHAMP, 
	include = c (GAMES, WINPERCENT, PAKE, PAKERANK, PASE, PASERANK, TOP2),
	label = list(
		GAMES ~ "Tournament Games Played" ,
		WINPERCENT ~ "Winning Percent", 
		PAKE ~ "Performance Against Komputer Expectations (PAKE)" ,
		PAKERANK ~ "PAKE Rank" ,
		PASE ~ "Performance Against Seed Expectations (PASE)",
		PASERANK ~ "PASE Rank",
		TOP2 ~ "Times Team was 1 or 2 Seed"
	),
	missing_text = "Missing"
) %>%
	add_overall()

table1

median_games <- inline_text(table1, variable = GAMES, column = "stat_0")

median_games_2 <- median(team_results$GAMES[team_results$CHAMP == 2])
```

The median number of tournament games played from 2008-2024 (excluding 2020) among all teams is `r median_games`.

Among teams that won two championships from 2008-2024, the median number of tournament games played was `r median_games_2`.

## Univariate Linear Regression

```{r}
#| label: tbl-regression
#| tbl-cap: "Regression of Number of Championships on 5 Characteristics"

tbl_uvregression(
	team_results,
	y = CHAMP,
	include = c(
		GAMES, WINPERCENT, 
		PAKERANK, PASERANK, TOP2
	),
	label = list(
		GAMES = "Tournament Games Played" ,
		WINPERCENT = "Winning Percent", 
		PAKE = "Performance Against Komputer Expectations (PAKE)" ,
		PAKERANK = "PAKE Rank" ,
		PASE = "Performance Against Seed Expectations (PASE)",
		PASERANK = "PASE Rank",
		TOP2 = "Times Team was 1 or 2 Seed"
	),
	method = lm
)

```

## Figures and Function

@fig-one shows the distribution of all team's performance against seed expectations (PASE).

```{r}
#| label: fig-one
#| fig-cap: "Histogram of All Team's Performance Against Seed Expectations"

hist(team_results$PASE)

```

```{r}

avgTOP2 <- function(x) {
	sum <- sum(team_results$TOP2[team_results$CHAMP == x])
	n <- length(team_results$TOP2[team_results$CHAMP == x])
	avg = sum/n
	return(avg)
}

avg_TOP2 <- c(avgTOP2(0), avgTOP2(1), avgTOP2(2), avgTOP2(3))

avg_TOP2_df <- data.frame(
CHAMP = c(0:3),
avg_TOP2
)



```

```{r}
#| label: fig-two
#| fig-cap: "Average Number of Top 2 Seeds Recieved by Number of Champtionships Won 2008-2024 (Excluding 2020)"

ggplot(avg_TOP2_df, aes (x = CHAMP, y = avg_TOP2)) + 
	geom_col(fill = "steelblue") +
  labs(
       x = "Championships",
       y = "Average Number of Top 2 Seeds Recieved") +
  theme_minimal()

avg_TOP2_2 <- avgTOP2(2)

```

Among teams with 2 championships between 2008-2004 (excluding 2020), the average number of top 2 seeds received in the same time period is `r avg_TOP2_2`.
